---
title: "Genomics I"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read sample genotypes data from ENSEMBLE

We downloaded genotype data from ENSEMBLE on the MXL Mexican Ancestry in LA, CA dataset

What proportion of this data are G/G, etc?

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
mxl
```

We want to look at the second column that contains the genotype information

```{r}
table(mxl$Genotype..forward.strand.)
```

## RNA-Seq result analysis for different genotypes of this SNP

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
expr
```

```{r}
expr[expr$geno == "G/G",]$exp
summary(expr[expr$geno == "G/G",]$exp)
```

```{r}
summary(expr[expr$geno == "A/G",]$exp)
```

```{r}
summary(expr[expr$geno == "A/A",]$exp)
```


```{r}
hist(expr[expr$geno == "G/G",]$exp) #not great for visualization
```

##Try a box plot

We will use the boxplot function and the input data will be **expr**. How do we draw a useful plot?

```{r}
boxplot(exp ~ geno, data=expr, notch=TRUE)
```

How many samples are we looking at?

```{r}
nrow(expr)
```



```{r}
boxplot(expr[expr$geno == "G/G",]$exp, expr[expr$geno == "A/G",]$exp, expr[expr$geno == "A/A",]$exp)
```



Overall, the pipeline we went through today: FASTQ -> QC -> Mapping -> Counting -> Differential expression 
















