---
title: "Structure based drug design"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and process starting structure

Here we download and clean up the HIV-Pr structure (pdb code: 1HSG) from the main PDB database. We will make a separate set of "protein-only" and "ligand only" PDB files.

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
file.name
```

We will use `read.pbd`, `atom.select`, and `write.pdb` functions to make our separate protein and ligand only pdb files

```{r}
hiv <- read.pdb("1hsg")
hiv
```


```{r}
protein <- atom.select(hiv, "protein", value = TRUE)
ligand <- atom.select(hiv, "ligand", value = TRUE)
protein
ligand
```

```{r}
write.pdb(protein, file = "1hsg_protein.pdb")
write.pdb(ligand, file = "MK_ligand.pdb")
#we made separate files out of the information we downloaded from pbd, files appeared in file folder
```

Then we use AutoDock to get a file with a charge and atom type column

## Read docking results

Read in the output of docking ans make a PDB file for viewing in VMD or PyMol

```{r}
res <-read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb") # now you can open that file in pymol
```
















