---
title: "Class8"
author: "Rimma Levina"
date: "2/5/2020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Lets try the `kmeans()` function in R to cluster some made up example data

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3)) #this makes up data, 1st = 30 points centered around -3, 2nd = 30 points centered around +3 --> makes a vector that has 60 points

x <- cbind(x=tmp, y=rev(tmp)) # cbind() binds together vectors, you are also setting the axes, where x is the first vector and y is the reverse (rev()) of the tmp vector, then you plot and see you have 2 groups that are clearly separate/clustered
 
plot(x)
```
Use the kmeans() function setting k to 2 and nstart=20 
Inspect/print the results 
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
         - cluster size?      
         - cluster assignment/membership?       
         - cluster center?

Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
km <- kmeans(x = x, centers = 2, nstart = 20) 
km
```


attributes(km) is a function to get the details of your data



```{r}
km$size
```

```{r}
km$cluster
```

```{r}
table(km$cluster)
```

Now plot colored by the kmeans cluster assignment, and add cluster centers as blue points
so find CENTERS, then add to exsiting plot

```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue", pch=16, cex=3)
```

## Hierarchical clustering in R

The `hclust()` function is the main hierarchical clustering method in R and it MUST BE PASSED THROUGH A MATRIX AS INPUT, not your raw data!!!!

```{r}
hc <- hclust( dist(x))  # this hopefully reveals patterns in your data
plot(hc)
```

```{r}
plot(hc)
abline(h=6, col="red", lty= 2)# makes a line above where you want to cut off the data
abline(h=3.5, col="blue", lty= 2)
```

```{r}
cutree(hc, h=6)  # lists all the points below that line you cut
```
You can call cutree with `k` for the number of groups you want, or with `h` for height

```{r}
cutree(hc, k=6) # having a lower cut off you see now you have 6 clusters
```

## Linkage

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd=0.3), ncol = 2), # c1
 matrix(rnorm(100, mean=1, sd=0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean=1, sd=0.3), # c3
 rnorm(50, mean=0, sd=0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

Q. Use the dist() for distance matrix, hclust(), plot() and cutree() functions to return 2 and 3 clusters 
Q. How does this compare to your known 'col' groups?

```{r}
hc <- hclust(dist(x))
plot(hc)
```

```{r}
grps2 <- cutree(hc, k=2)
grps3 <- cutree(hc, k=3) #these are membership vectors
```


```{r}
table(grps2)
table(grps3)
```

```{r}
plot(x, col=grps3) # color by the group distribution
```

```{r}
table(grps3, col) #this is a cross table, using col from line 102, saying for group 1 all 42 points come from cluster 1, for group 2 most points come from cluster 3, for group 3 most points come from cluster 2 - so only group 1 is mostly right
```



# Principle Component Analysis (PCA)

The main function in R for PCA is called `prcomp()`. Here we will use PCA to examine the funny food that folks eat in the UK and N. Ireland.

Import the CSV file first:

```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x

# Notice that X is the lable for the first column, but you want that to be the row names. You can fix that in the import function, with row.names=1

```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```


```{r}
pairs(x, col=rainbow(10), pch=16)  #Makes a matrix of scatterplots
```

#PCA to the rescue!

```{r}
pca <-prcomp(t(x)) #taking the transpose of x makes the rows the column names
```

```{r}
summary(pca)
```

```{r}
attributes(pca)
```

```{r}
plot( pca$x[,1], pca$x[,2] )
text(pca$x[,1], pca$x[,2], colnames(x), col=c("black", "red", "blue", "darkgreen"))
```










