---
title: "Structural Bioinformatics (pt.1)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats #when you visualize it from the environment tab, you might want to rename the columns
```

Q. What percentage of structures are solved by X-Ray and Electron Microscopy?

```{r}
ans <- round((stats$Total / sum(stats$Total))*100, 2) # gives you the calculation for every row
names(ans) <- rownames(stats)
ans
```

Q. What proportion of structures are protein?

```{r}
prop <- round(sum(stats$Proteins/sum(stats$Total)), 2)
prop
```


## Section 2: Using bio3d package to analyze protein structure data

Q. Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

Use the bio3d package from CRAN. 

```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

Adenylate kinase structures from the PDB for analysis

```{r}
library(bio3d)
aa <- get.seq("1ake_A")
aa
```


```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A',
                 '3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A',
                 '6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A',
                 '4PZL_A')
```

Download and parse these structures

```{r}
library(bio3d)
#Download related PDB files
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
```


Align and superimpose structures

```{r}
# Align related PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile = "msa")
```

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```


```{r}
library(bio3d.view)
library(rgl)
view(pdbs) 

# rglwidget(elementId=" tmp") cant make a markdown file with this because it procudes an html output
```

```{r}
anno <- pdb.annotate(ids)
unique(anno$source)
```

```{r}
pc.xray <- pca(pdbs)
plot(pc.xray)
```
```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

```{r}
#Plotting results with ggplot2
library(ggplot2)
library(ggrepel)

df <- data.frame(x=pc.xray$z[,1], y=pc.xray$z[,2])
col <- as.factor(grps.rd)

p <- ggplot(df, aes(x, y)) +
 geom_point(aes(col=col), size=2) +
 xlab("PC1") +
 ylab("PC2") +
 scale_color_discrete(name="Clusters") +
 geom_text_repel(aes(label=ids))
p

```

