---
title: "Network Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coronavirus (and HW)

Here we analyze infection data for the 2019 novel Coronavirus COVID-19 (2019-nCoV) epidemic. The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus repository.

A CSV file is available here https://github.com/RamiKrispin/coronavirus-csv

```{r}
url <- "https://tinyurl.com/COVID-2019"
virus <- read.csv(url)

tail(virus)
```

> Q1. How many total cases around the world?

```{r}
total <- sum(virus$cases)
total
```

> Q2. How many deaths linked to infected cases have there been?

```{r}
table(virus$type)
inds <- virus$type == "death"
virus[inds, ]
```

```{r}
totaldeaths <- sum(virus[inds, "cases"])
totaldeaths
```


> Q3. What is the overall dealth rate?

```{r}
round(totaldeaths/total * 100, 2)
```


> Q4. What is the death rate in Mainland China?

```{r}
totalchina <- virus$Country.Region == "Mainland China"
tc<- sum(virus[totalchina, "cases"])

chinadeath <- virus$type == "death" & virus$Country.Region == "Mainland China"
cd<- sum(virus[chinadeath, "cases" ])

round(cd/tc * 100, 2)

```


> Q5. What is the death rate in Italy, Iran, and the US?

```{r}
#Italy

totalitaly <- virus$Country.Region == "Italy"
ti<- sum(virus[totalitaly, "cases"])

italydeath <- virus$type == "death" & virus$Country.Region == "Italy"
id<- sum(virus[italydeath, "cases" ])

round(id/ti * 100, 2)
```

```{r}
#Iran

totaliran <- virus$Country.Region == "Iran"
tir<- sum(virus[totaliran, "cases"])

irandeath <- virus$type == "death" & virus$Country.Region == "Iran"
ird<- sum(virus[irandeath, "cases" ])

round(ird/tir * 100, 2)
```

```{r}
#US

totalUS <- virus$Country.Region == "US"
tUS<- sum(virus[totalUS, "cases"])

USdeath <- virus$type == "death" & virus$Country.Region == "US"
USd<- sum(virus[USdeath, "cases" ])

round(USd/tUS * 100, 2)

```















